---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jueun Kang
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#### Quarto workspace setup and reading csv's into quarto ####
library(tidyverse)
library(palmerpenguins)
library(haven)
library(here)

# respondents' degree obtained 
clean_data_degree <- read_csv("~/Documents/environment_education/inputs/data/degrees_clean.csv")

# respondents' willingness to donate towards protecting the environment
clean_data_willingness <- read_csv("~/Documents/environment_education/inputs/data/willingness_clean.csv")

# respondents who donated to an environmental group
clean_data_donation <- read_csv("~/Documents/environment_education/inputs/data/donation_clean.csv")

```


# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point. 



# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

clean_data_degree <-
  clean_data_degree |>
  mutate(
    degree = 
      recode(
        degree,
        "0" = "< High School",
        "1" = "High School",
        "2" = "Junior College",
        "3" = "Bachelors",
        "4" = "Graduate"
      )
  )

clean_data_degree |>
  ggplot(aes(x = degree)) +
  geom_bar(fill = "slateblue1") +
  theme_minimal() +
  labs(x = "Degree Obtained",
       y = "Number of Participants",
       title = "Degree obtained by participant")
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


